Reduced GW invariants from cuspidal curves

### 2017-12-08 draft.tex
missing only the last step of the proof

We are now going to construct the cosection as in \cite[\S\S3.2-3.4]{CLpfields}. There is a morphism of vector bundles on $\hP$ induced by tensoring of line bundles:
\[ h_1\colon\rm{Vb}(\mathcal L_{\hP}^{\oplus 5}\oplus\mathcal P_{\hP})\to \rm{Vb}(\omega_{\hat\pi_{\hP}}),\quad h_1(z,p)=p\w(z_0,\ldots,z_4)\]
By differentiating it and pulling it back along the universal evaluation
\bcd
& \rm{Vb}(\mathcal L_{\hP}^{\oplus 5})\setminus\{0\}\oplus\rm{Vb}(\mathcal P_{\hP})\ar[d] \\
\hat \cC_{\pazocal P}\ar[ur,bend left,"\mathfrak e"]\ar[d,"\hat\pi_{\pazocal P}"]\ar[r] & \hat\cC_{\hP}\ar[d,"\hat\pi_{\hP}"] \\
\pazocal P \ar[r] & \rm \hP
\ecd
(here and in future $\oplus$ at the level of geometric vector bundles means the fiber product over the base) we obtain a cosection of the relative obstruction sheaf
\begin{equation*}\label{eqn:cosection}
 \begin{split}
 \sigma_1\colon\rm{Ob}_{\pazocal P/\hP}=\R^1\hat\pi_{\pazocal P,*}(\mathcal L_{\pazocal P}^{\oplus 5}\oplus\mathcal P_{\pazocal P})\to \R^1\hat\pi_{\pazocal P,*}(\omega_{\hat\pi_{\pazocal P}})\simeq \OO_{\pazocal P} \\
 \sigma_{1|(u,\psi)}(x,p)=p\w(u)+\psi\sum_{i=0}^4\partial_i\w(u)x_i
\end{split}
\end{equation*}

@@

The key point is that $\mathbb E\rvert_{\tZ^{p,\rm{gst}}}$ has locally free $h^0$ \emph{and} $h^1$, as can be seen from the equations. Then, when we pick a resolution by locally free sheaves $\mathbb E=[F_0\xrightarrow{d} F_1]$, the image sheaf $d(F_0)$, which is the kernel of $F_1\to h^1(\mathbb E\rvert_{\tZ^{p,\rm{gst}}})$, is really a sub-vector bundle of $F_1$. Then the projections:
\[
 \beta\colon F_1\to h^1/h^0(\mathbb E) \quad \text{and} \quad \beta'\colon F_1\to \tilde{V}:=\R^1\hat{\pi}_*(\hat{\mathcal L}^{\oplus 5}\oplus\mathcal P)
\]
are both flat. The cone stack $\mathfrak C^{\rm{gst}}$ can be descended to a cone $C^{\rm{gst}}$ by means of taking the quotient of $\beta^{-1}\mathfrak C^{\rm{gst}}$ by the free action of $d(F_0)$; $C^{\rm{gst}}$ should then be thought of as the coarse moduli of $\mathfrak C^{\rm{gst}}$. Recall that the cosection $\sigma$ is induced by a cosection $\sigma_1$ of $\tilde V$ (see Eq. \eqref{eqn:cosection}). It follows from the commutativity of localised Gysin pullback with flat pullback that:
\[
 0^!_{h^1/h^0(\mathbb E),\sigma}(\mathfrak C^{\rm{gst}})= 0^!_{\tilde V,\sigma_1}(C^{\rm{gst}})
\]
See \cite[Proposition 6.3]{CLpfields}.

They then introduce the notion of \emph{homogeneity} for substacks of $\tilde V$: write $\tilde V=\tilde V_1\oplus\tilde V_2$ with $\tilde V_1=\R^1\hat{\pi}_*(\hat{\mathcal L}^{\oplus 5})$ and $\tilde V_2=\R^1\hat{\pi}_*(\mathcal P)$, and $\gamma\colon \tZ^{p,\rm{gst}}\to\tZ^{\rm{gst}}$ the projection. Since $\tZ^{p,\rm{gst}}$ is the total space of a line bundle over $\tZ^{\rm{gst}}$, it comes with a natural $\Gm$-action on the fibers of $\gamma$, and since $\tilde V_i=\gamma^*V_i$ for the corresponding vector bundles $V_i$ on $\tZ^{\rm{gst}}$, the total space of $\tilde V_i$ can be endowed with a $\Gm$-action that makes the projection to $\tZ^{p,\rm{gst}}$ equivariant. A closed substack of $\tilde V$ is $0$-homogeneous if it is the pullback of a closed substack of $V$ along $\gamma$. In facts the $\tilde V_i$'s can be endowed with different $\Gm$ actions that make the projection to $\tZ^{p,\rm{gst}}$ equivariant by twisting with two characters of $\Gm$, one for each $\tilde V_i$; this defines the concept of $(l_1,l_2)$-homogeneous substacks of $\tilde V$. Here is how we are going to use the homogeneity:

@@

\begin{comment}{
We may notice that $(\Z^p)^{\text{main},\circ}\to\XP$ is smooth onto its image, which is an open substack; furthermore on this open the relative tangent is represented by $\hat\pi_*\hL^{\oplus 5}$, and $\R^1\hat\pi_*\hL^{\oplus 5}=\hat\pi_*(\hat\omega\otimes\hL^{\otimes -5})=0$. Let $E^\bullet=\R^\bullet\hat\pi_*(\hL^{\oplus 5}\oplus \hat\omega\otimes\hL^{\otimes -5})$ denote the dual perfect obstruction theory of $\Z^p\to \XP$. Then the intrinsic normal cone of $\Z^p\to \XP$ splits into:
\begin{itemize}
\item the closure of the zero section of $h^1/h^0(E^\bullet)_{|\text{main},\circ}$, which we shall name $\mathfrak C^{\text{main}}$;
\item other components supported on the boundary, which we should accordingly name $\mathfrak C^k$ ($k\geq 2$).
\end{itemize}}
\end{comment}
\bibliographystyle{alpha}
\bibliography{cusp}

>>>>

We are now going to construct the cosection as in \cite[\S\S3.2-3.4]{CLpfields}. There is a morphism of vector bundles on $\hP$ induced by tensoring of line bundles:
\[ h_1\colon\rm{Vb}(\mathcal L_{\hP}^{\oplus 5}\oplus\mathcal P_{\hP})\to \rm{Vb}(\omega_{\hat\pi_{\hP}}),\quad h_1(z,p)=p\w(z_0,\ldots,z_4)\]
By differentiating it and pulling it back along the universal evaluation
\bcd
& \rm{Vb}(\mathcal L_{\hP}^{\oplus 5})\setminus\{0\}\oplus\rm{Vb}(\mathcal P_{\hP})\ar[d] \\
\hat \cC_{\pazocal P}\ar[ur,bend left,"\mathfrak e"]\ar[d,"\hat\pi_{\pazocal P}"]\ar[r] & \hat\cC_{\hP}\ar[d,"\hat\pi_{\hP}"] \\
\pazocal P \ar[r] & \rm \hP
\ecd
(here and in future $\oplus$ at the level of geometric vector bundles means the fiber product over the base) we obtain a cosection of the relative obstruction sheaf
\begin{equation}\label{eqn:cosection}
 \begin{split}
 \sigma_1\colon\rm{Ob}_{\pazocal P/\hP}=\R^1\hat\pi_{\pazocal P,*}(\mathcal L_{\pazocal P}^{\oplus 5}\oplus\mathcal P_{\pazocal P})\to \R^1\hat\pi_{\pazocal P,*}(\omega_{\hat\pi_{\pazocal P}})\simeq \OO_{\pazocal P} \\
 \sigma_{1|(u,\psi)}(x,p)=p\w(u)+\psi\sum_{i=0}^4\partial_i\w(u)x_i
\end{split}
\end{equation}

@@

The key point is that $\mathbb E\rvert_{\tZ^{p,\rm{gst}}}$ has locally free $h^0$ \emph{and} $h^1$, as can be seen from the equations. Then, when we pick a resolution by locally free sheaves $\mathbb E=[F_0\xrightarrow{d} F_1]$, the image sheaf $d(F_0)$, which is the kernel of $F_1\to h^1(\mathbb E\rvert_{\tZ^{p,\rm{gst}}})$, is really a sub-vector bundle of $F_1$. Then the projections:
\[
 \beta\colon F_1\to h^1/h^0(\mathbb E) \quad \text{and} \quad \beta'\colon F_1\to \tilde{V}:=\R^1\hat{\pi}_*(\hat{\mathcal L}^{\oplus 5}\oplus\mathcal P)
\]
are both flat. The cone stack $\mathfrak C^{\rm{gst}}$ can be descended to a cone $C^{\rm{gst}}$ by means of taking the quotient of $\beta^{-1}\mathfrak C^{\rm{gst}}$ by the free action of $d(F_0)$; $C^{\rm{gst}}$ should then be thought of as the coarse moduli of $\mathfrak C^{\rm{gst}}$. Recall that the cosection $\sigma$ is induced by a cosection $\sigma_1$ of $\tilde V$ (see Eq. \eqref{eqn:cosection}). 
\marginpar{rewrite what they are}It follows from the commutativity of localised Gysin pullback with flat pullback that:
\[
 0^!_{h^1/h^0(\mathbb E),\sigma}(\mathfrak C^{\rm{gst}})= 0^!_{\tilde V,\sigma_1}(C^{\rm{gst}})
\]
See \cite[Proposition 6.3]{CLpfields}.

They then introduce the notion of \emph{homogeneity} for substacks of $\tilde V$: write 
\[\tilde V=\tilde V_1\oplus\tilde V_2\;\;\text{ with} \;\;\tilde V_1=\R^1\hat{\pi}_*(\hat{\mathcal L}^{\oplus 5})\;\; \text{and}\;\; \tilde V_2=\R^1\hat{\pi}_*(\mathcal P),\]
 and $\gamma\colon \tZ^{p,\rm{gst}}\to\tZ^{\rm{gst}}$ the projection. Since $\tZ^{p,\rm{gst}}$ is the total space of a line bundle over $\tZ^{\rm{gst}}$, it comes with a natural $\Gm$-action on the fibers of $\gamma$. Moreover, since 
 \[\tilde V_i=\gamma^*V_i\]
 \marginpar{say few words, that is not the definition of $\tilde V_i$ but a property (Lemma 6.1)}
 for the corresponding vector bundles $V_i$ on $\tZ^{\rm{gst}}$, the total space of $\tilde V_i$ can be endowed with a $\Gm$-action that makes the projection to $\tZ^{p,\rm{gst}}$ equivariant. A closed substack of $\tilde V$ is $0$-homogeneous if it is the pullback of a closed substack of $V$ along $\gamma$. In facts the $\tilde V_i$'s can be endowed with different $\Gm$ actions that make the projection to $\tZ^{p,\rm{gst}}$ equivariant by twisting with two characters of $\Gm$, one for each $\tilde V_i$; this defines the concept of $(l_1,l_2)$-homogeneous substacks of $\tilde V$. Here is how we are going to use the homogeneity:

@@

\begin{comment}{
We may notice that $(\Z^p)^{\text{main},\circ}\to\XP$ is smooth onto its image, which is an open substack; furthermore on this open the relative tangent is represented by $\hat\pi_*\hL^{\oplus 5}$, and $\R^1\hat\pi_*\hL^{\oplus 5}=\hat\pi_*(\hat\omega\otimes\hL^{\otimes -5})=0$. Let $E^\bullet=\R^\bullet\hat\pi_*(\hL^{\oplus 5}\oplus \hat\omega\otimes\hL^{\otimes -5})$ denote the dual perfect obstruction theory of $\Z^p\to \XP$. Then the intrinsic normal cone of $\Z^p\to \XP$ splits into:
\begin{itemize}
\item the closure of the zero section of $h^1/h^0(E^\bullet)_{|\text{main},\circ}$, which we shall name $\mathfrak C^{\text{main}}$;
\item other components supported on the boundary, which we should accordingly name $\mathfrak C^k$ ($k\geq 2$).
\end{itemize}}
\end{comment}
\subsection*{Step II: from cosection localized to standard Gysin map}
Recall that when we are working on a proper DM stack $\pazocal M$ endowed with a perfect obstruction theory and admitting a cosection $\sigma$, then the cosection localized Gysin map can be reduced to the ordinary Gysin map, i.e.
\[\iota_*\circ 0^{!}_{E,\sigma}=0_E^{!}\circ \tilde{\iota}\;\;\;\text{where}\;\;\iota\colon D(\sigma)\hookrightarrow\pazocal M,\;\;\tilde{\iota}\colon E(\sigma)\hookrightarrow E\]
and $E$ is the vector bundle containing the cone.

Then to compute $0^!_{\tilde V,\sigma_1}(C^{\rm{gst}})$ we plan to compactify $\tZ^{p,\rm{gst}},$ extend the cone, the vector bundle and the section over it and then make use of the fact we just recalled.

Since $\tZ^{p,\rm{gst}}\cong\operatorname{Tot}_{\tZ^{\rm{gst}}}\left(\R^1\hat{\pi}_*\mathcal P\right)$ we compactify considering
\[\overline{\gamma}\colon \overline{\tZ^{p,\rm{gst}}}:=\PP\left(\R^1\hat{\pi}_*\mathcal P\oplus \OO_{\tZ^{\rm{gst}}}\right)\to \tZ^{\rm{gst}}\]
and extend $\tilde{V}_1,\tilde{V}_2$ via:
\[\tilde{V}^{\rm{cpt}}_1=\overline{\gamma}V_1(-D_{\infty}),\qquad \tilde{V}^{\rm{cpt}}_2=\overline{\gamma}V_2.\]
Finally, let us extend the cosection as well. From the cosection $\sigma$ defined by \ref{eqn:cosection} we clearly get a cosection on $\Z^p$ by pull-back and again a 
\[\widetilde{\sigma}\colon \pazocal Ob_{\tZp}\to\OO_{\tZp} \]
 on the blow-up. Let us denote by $\widetilde{\sigma}_{\rm{gst}}$ its restriction to $\tZ^{\rm{gst}}.$
 Writing down what $\widetilde{\sigma}_{\rm{gst}}$ looks like on points\cite[Lemma~6.2]{CL}, we see that it extend to a homomorphism of vector bundles:
 \[\overline{\sigma}\colon\tilde{V}^{\rm{cpt}}=\tilde{V}^{\rm{cpt}}_1\oplus \tilde{V}^{\rm{cpt}}_2\to \overline{\tZ^{p,\rm{gst}}}.\]
 \begin{prop}
 Let $\iota_{!}\colon Z_*(\tilde{V}(\widetilde{\sigma}_{\rm{gst}}))\to Z_*(\tilde{V}^{\rm{cpt}})$ be defined by 
 $\iota_{!}[C]=[\overline{C}].$ And $i\colon D(\widetilde{\sigma}_{\rm{gst}})\to \tilde{Z}_{\rm{gst}}$ the inclusion. Then
 \[\overline{\gamma}_*\circ 0^{!}_{\tilde{V}^{\rm{cpt}}}\circ\iota_{!}=i\circ 0^{!}_{\widetilde{\sigma}_{\rm{gst}},\rm{loc}}\colon  Z_*(\tilde{V}(\widetilde{\sigma}_{\rm{gst}}))\to A_*(\tilde{Z}_{\rm{gst}}).\]
 \end{prop}
 This follows essentially from the property of cosection localized cycles recalled at the beginning of the section; see \cite[Proposition~6.4]{CL} for full details.
 \subsection*{Step III: reduction of the support of the cone}
We now explain a key technical lemma which will enable us to show that $C^{\rm{gst}}$ pushes forward to zero under a suitably defined morphism. It is basically reducing the support of $C^{\rm{gst}}\cap 0\oplus\tilde V_2$ to a manageable substack of $\tilde V_2$, that is the union of the zero-section (i.e. $\tZp$) and a sub-line bundle of $\tilde V_2$ supported on $\tilde \Delta=\tZ^{p,\rm{main}}\cap\tZ^{p,\rm{gst}}$. Even better, using the homogeneity we can show that such a line bundle comes from $\Delta=\tZ^{\rm{main}}\cap\tZ^{\rm{gst}}$. This is \cite[Proposition 7.1]{CLpfields}.
\begin{lem}
 There is a sub-line bundle $F$ of $V_2\rvert_\Delta$ such that:
 \[
  C^{\rm{gst}}\cap 0\oplus\tilde V_2\subseteq 0_{\tilde V_2}\cup \tilde F:= \tZ^{p,\rm{gst}}\cup \gamma^* F
 \]
\end{lem}
First they use the fact that there is a triple of compatible obstruction theories for the triangle:
\bcd
\tZp \ar[rr]\ar[dr,"\gamma"] & & \tZ \ar[dl] \\
& \XP &
\ecd
such that their restrictions to $\tZ^{p,\rm{gst}}$ have locally free $h^0$ and $h^1$. To the effect that by taking $h^1$ of the dual obstruction theories we obtain a commutative diagram:
\bcd
h^1(\mathbb L_{\tZp/\tZ}^\vee\rvert_{\tZ^{p,\rm{gst}}}) \ar[r]\ar[d] & h^1(\mathbb L_{\tZp/\XP}^\vee\rvert_{\tZ^{p,\rm{gst}}})\ar[r]\ar[d] &h^1(\gamma^*\mathbb L_{\tZ/\XP}^\vee\rvert_{\tZ^{p,\rm{gst}}})\ar[d] \\
\tilde V_2\ar[r,"i_2"] & \tilde V_1\oplus\tilde V_2 \ar[r,"\pr_1"] & \tilde V_1
\ecd
The vertical arrows are injective by the definition of an obstruction theory, and the bottom triangle is exact. Notice that $0\oplus\tilde V_2$ is precisely the kernel of $\pr_1$. It follows that, in order to understand the support of $C^{\rm{gst}}\cap 0\oplus\tilde V_2$, it is enough to study that of $N$, where $N$ is the coarse moduli cone of $\mathfrak C_{\tZp/\tZ}$, living in the upper left corner of the above diagram.

This is an easier task, since we know that $\tZp/\tZ$ is a line bundle on $\tZ^{\rm{gst}}$ and an isomorphism on $\tZ^{\rm{main},\circ}$. Hence we can always find a local chart $S\to \tZ$ such that the following diagram holds:
\bcd
\tZp\ar[d] & T\ar[l]\ar[ld,phantom,"\Box"]\ar[d]\ar[r,hook,"V(\zeta t)"] & S\times\Aaff^1_t\ar[ld] \\
\tZ & S\ar[l,"\acute{e}t"] &
\ecd
where $\zeta$ is a local equation for the boundary. Then $\tau^{\geq-1}\mathbb L_{\tZp/\tZ}\rvert_T=[I/I^2\xrightarrow{\delta}\Omega_{\Aaff^1_S/S}]$; $I$ is generated by $\zeta t$, whose image under $\delta$ is $\zeta \rm{d}t$, which restricts to $0$ on $\tZ^{p,\rm{gst}}\times_{\tZp}T=\{\zeta=0\}$. So the action is trivial, and in fact the restriction of the coarse moduli cone is precisely $\Spec_{T^{\rm{gst}}}\operatorname{Sym}^{\bullet} I/I^2$, which is a line bundle supported on $\tilde \Delta\times_{\tZp}T$ and trivial otherwise. By gluing different charts we get the line bundle $\tilde F$ on $\tilde \Delta$.

The last part of the statement, namely that $\tilde F$ descends to a line bundle $F$ on $\Delta$ is proved by homogeneity: the normal cone of $\tZp/\tZ$ is homogeneous with respect to the $\Gm$-action with character $1$ on the fibers of $\tilde V_2\to \tZ^{p,\rm{gst}}$, but being a cone it is $0$-homogeneous as well (see the above discussion of homogeneity), so it is $\gamma^*F$ for some line bundle $F$ on $\Delta\subseteq \tZ$.

\subsection*{Step IV: pushing forward to zero} Recall from the previous sections that we need to show that the degree of the following class is $0$:
\[
 0_{\overline{V}}^![\overline C^{\rm{gst}}]=0_{\overline V_2}^!(0_{\overline{V_1}}^![\overline C^{\rm{gst}}])=0_{\overline V_2}^!([N_{\overline C^{\rm{gst}}\cap (0\oplus \overline V_2)}\overline C^{\rm{gst}}])
\]
\bcd
? \ar[r]\ar[d] & X\ar[d] & \\
\overline C^{\rm{gst}}\cap (0\oplus \overline V_2) \ar[r]\ar[d] & V_2\ar[r]\ar[d] & X\ar[d] \\
\overline C^{\rm{gst}} \ar[r] & \overline V\ar[r] & \overline V_1
\ecd
We know from the previous section that the last cycle can be split into two parts, one (call it $N_1$) supported on a sub-line bundle of $\overline V_2$ on $\tilde\Delta$, the other one (call it $N_2$) suypported on the zero section of $\overline V_2$.

\begin{lem}
 Both $N_1$ and $N_2$ are $5d+1$-dimensional cycles, and $\deg(0_{\overline V_2}^![N_i])=0$ for $i=1,2$.
\end{lem}
\begin{proof}
 Compare with \cite[Lemma 8.1]{CLpfields}. The dimension of $\tZ^{\rm{gst}}$ is $5d+3$, being locally a $5(d+1)$ vector bundle over a dimension $-2$ stack; so $\tZ^{p,\rm{gst}}$, which is a line bundle on the former, has dimension $5d+4$. The coarse moduli cone has then dimension $5d+6$, as can be argued from \ref{lem:open_cones}. $\overline V_1\rvert_{\tZ^{p,\rm{gst}}}$ has rank $5$, so $0_{\overline{V_1}}^![\overline C^{\rm{gst}}]$ is a cycle of dimension $5d+1$.
\end{proof}

\bibliographystyle{alpha}
\bibliography{cusp}
